The Pair Distribution Function of the Inhomogeneous Hard-Sphere Fluid

### 2013-11-18 papers/pair-correlation/paper.tex
edit fit section of pair-correlation paper Tested on collins. Took .

The function has the form:
\begin{align}
  g(|\rr|) &= 1 + g_{\sigma}e^{-x_0(r-\sigma)} + x_1g_{\sigma}\sin(x_2(r-\sigma))e^{-x_3(r-\sigma)}\\
  &~~-x_4g_{\sigma}^2\sin(x_5(r-\sigma))e^(-x_6(r-\sigma))
\end{align}
The finalalized form of our pair correlation function is:
\begin{align}
  g^{(2)}(\rr_1,\rr_2) = \frac{g(|r_{12}|, g_\sigma(\rr_1)) + g(|r_{12}|, g_\sigma(\rr_2))}{2}
\end{align}

%% gsigma = gsigconcatenated[i] - 1
%%         h0 = gsigma # was x[0]*gsig
%%         f0 = numpy.exp(-x[0]*rconcatenated[i])
%%         h1 = x[1]*gsigma
%%         f1 = numpy.sin(x[2]*rconcatenated[i]) * numpy.exp(-x[3]*rconcatenated[i])
%%         h2 = -x[4]*gsigma**(2)
%%         f2 = numpy.sin(x[5]*rconcatenated[i]) * numpy.exp(-x[6]*rconcatenated[i])
%%         g[i] = 1 + h0*f0 + h1*f1 + h2*f2
%%     return g

%% The homogeneoues limit $g_{sigm}$ can be calculated using the well
%% known and very successful Carnahan Starling equation while the
%% homogeneous limit radial distribution function can be calculated using
%% Monte-Carlo methods.  We used these calculations in a least squares
%% fit along the data points at increasing radial distances in order to
%% fit the perameters of our more general pair correlation function.  Our
%% function will therefore approach the most accurate solution possible
%% in the homogeneous limit.  Created a function that has the form of an
%% expansion in powers of $g_{\sigma}$.  $g_{sigm}$ in the homogeneous
%% limit can be calculated using the well known and very successful
%% Carnahan Starling equation.  has a well known and Perameters in the
%% expansion where fit
\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figs/ghs-g}% \\
  %\includegraphics[width=\columnwidth]{figs/ghs-g-ghs}
  \caption{Plot of the radial distribution function, with our fit.}\label{fig:radial-distribution}
\end{figure}
\subsection{Separable fit to the radial distribution function}

\jeffsays{I'm confused by this next part.  Are we just saying that
  instead of using our published $g_{\sigma}$ we could have used a
  Carnahan Starling derived $g_{\sigma}$? Don't understand why that
  would add an unnecessary step in using g(r).}

We have constructed a separable fit to the radial distribution
function $g(r, g_\sigma)$, which is a function of the radius $r$ and
its value at contact $g_\sigma$.  We could convert this to a more
conventional function of the packing fraction $\eta$ by inverting the
Carnahan-Starling formula for $g_\sigma$
\begin{equation}
  g_\sigma = \frac{1-\tfrac{\eta}{2}}{(1-\eta)^3}
\end{equation}
but that would add an unnecessary step in the use of $g(r)$, since the
averaged value of $g_\sigma$ is what we can directly compute for the
inhomogeneous system~\cite{schulte2012using}.

We construct $g(r, g_\sigma)$ using several constraints.  Naturally,
\begin{equation}
  g(\sigma, g_\sigma) = g_\sigma.
\end{equation}
In addition, we know that
\begin{align}
  1 + n\int\left[g(r)-1\right]d\rr &= nkT\chi_T
\end{align}
\begin{equation}
  \begin{split}
    \chi_T  &= -\frac{1}{V}\left(\frac{\partial V}{\partial
      p}\right)_T \\
    &= - \left(V\left(\frac{\partial p}{\partial
      V}\right)_T\right)^{-1}
  \end{split}
\end{equation}
\begin{equation}
    p = nkT\frac{1+\eta+\eta^2-\eta^3}{(1-\eta)^3}
\end{equation}
\begin{equation}
    \left(\frac{\partial p}{\partial V}\right)_T = \frac{\partial p}{\partial\eta}\frac{\partial\eta}{\partial V} + \frac{\partial p}{\partial n}\frac{\partial n}{\partial V}
\end{equation}
\begin{equation}
    n = \frac{N}{V}\Longrightarrow \frac{\partial n}{\partial V} = -\frac{N}{V^2} = -\frac{n}{V}
\end{equation}
\begin{equation}
    \eta = \frac{4\pi R^3N}{3V}\Longrightarrow \frac{d\eta}{dV} = -\frac{\eta}{V}
\end{equation}
\begin{equation}
  \begin{split}
    \left(\frac{\partial p}{\partial V}\right)_T &= -nkT\frac{\eta}{V}\left[\frac{1 + 2\eta - 3\eta^2}{(1-\eta)^3} + 3\frac{1 + \eta + \eta^2 - \eta^3}{(1-\eta)^4}\right]\\
    &\quad  - \frac{nkT}{V}\left[\frac{1+\eta+\eta^2-\eta^3}{(1-\eta)^3}\right]\\
    &= -nkT\frac{\eta}{V}\left[\frac{(1+2\eta-3\eta^2)(1-\eta) + 3 + 3\eta + 3\eta^2 - 3\eta^3}{(1-\eta)^4}\right]\\
    &\quad  - \frac{nkT}{V}\left[\frac{(1+\eta+\eta^2-\eta^3)(1-\eta)}{(1-\eta)^4}\right]\\
    &= -\frac{nkT}{V}\left[\frac{4\eta + 4\eta^2 - 2\eta^3}{(1-\eta)^4} + \frac{1 - 2\eta^3 + \eta^4}{(1-\eta)^4}\right]\\
  \end{split}
\end{equation}
\begin{equation}
    nkT\chi_T = \frac{(1-\eta)^4}{1 + 4\eta + 4\eta^2 - 4\eta^3 + \eta^4}
  %\int (g(r)-1)d\rr &= \frac{\chi_T}{kT} \\
  %&= \eta \eta \cdots \fixme{fill in, double-check}
\end{equation}
where $\chi_T$ is the isothermal compressibility, for which we use the
Carnahan-Starling equation of state.

An additional constraint that we use sets the slope of $g(r,
g_\sigma)$ at $r=\sigma$.  \fixme{explain this}

Given these constraints, we use a functional form of
\begin{equation}
  g(r,g_\sigma) = 1 + A e^{-a(r-\sigma)}
  + B (r-\sigma) e^{-b(r-\sigma)} + C (r-\sigma)^2 e^{-c(r-\sigma)}
\end{equation}
where $A$, $B$, and $C$ are amplitudes determined by the above
constraints, and $a$, $b$, and $c$ are parameters fit to Monte Carlo
simulations of the hard-sphere radial distribution function.

\fixme{add paragraphs and equations with $A$, $B$ and $C$}

\fixme{add table with parameters}

Figure~\ref{fig:radial-distribution} shows our function
$g(r,g_\sigma)$ plotted against the $g(r)$ found from simulation.

>>>>

The function has the form:
\begin{multline}
  g(|\rr|) = 1 + g_{\sigma}e^{-x_0(r-\sigma)} \\+ x_1g_{\sigma}\sin(x_2(r-\sigma))e^{-x_3(r-\sigma)}\\
  -x_4g_{\sigma}^2\sin(x_5(r-\sigma))e^(-x_6(r-\sigma))
\end{multline}
The finalalized form of our pair correlation function is:
\begin{align}
  g^{(2)}(\rr_1,\rr_2) = \frac{g(|r_{12}|, g_\sigma(\rr_1)) + g(|r_{12}|, g_\sigma(\rr_2))}{2}
\end{align}

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figs/ghs-g2}% \\
  %\includegraphics[width=\columnwidth]{figs/ghs-g-ghs}
  \caption{Plot of the radial distribution function, with our fit.}\label{fig:radial-distribution}
\end{figure}
\subsection{Separable fit to the radial distribution function}

In contrast to the previous work, which are based on the homogeneous
radial distribution function $g(r,\eta)$ expressed as a function of
distance and packing fraction, we have constructed our separable
approximation for the radial distribution function as $g(r,
g_\sigma)$, which is a function of the radius $r$ and its value at
contact $g_\sigma$.  One can convert between these two representation
using the Carnahan-Starling formula for $g_\sigma(\eta)$
\begin{equation}
  g_\sigma = \frac{1-\tfrac{\eta}{2}}{(1-\eta)^3}.
\end{equation}
The form $g(r, g_\sigma)$ is more convenient for treating the
inhomogeneous case, because we can directly compute the averaged value
of $g_\sigma$ for an inhomogeneous system~\cite{schulte2012using}.

We construct $g(r, g_\sigma)$ to satisfy several constraints.  Naturally,
\begin{equation}
  g(\sigma, g_\sigma) = g_\sigma.
\end{equation}
A second constraint applies to the integral of the total correlation
function
\begin{align}
  1 + n\int h(r)d\rr &= nkT\chi_T \label{eq:total-constraint}
\end{align}
where $h(r) = g(r) - 1$ is the total correlation function.
A final constraint is that the slope of the radial distribution
function at contact is given to a very good approximation by
\begin{align}
  g'(\sigma) \approx - h_\sigma g_\sigma
\end{align}
where $h_\sigma = g_\sigma - 1$ is the total correlation function.
Under these two constraints, we applied a least-squares fit to
simulation data for the radial distribution function at packing
fractions from 0.05 to 0.4 by steps of 0.05.  Our fitted radial
distribution function is given by
\newcommand\kappaa{\kappa_0}
\newcommand\kappab{\kappa_1}
\newcommand\kappac{\kappa_2}
\begin{multline}
  h(r,g_\sigma) = h_\sigma e^{-\kappaa z} \\
  + h_\sigma(\kappaa - g_\sigma) ze^{-\kappab z} \\
  + \frac{\frac{(1-\eta)^4}{1+4\eta + 4\eta^2 - 4\eta^3 +
      \eta^4}\frac{1}{n}
    - I_0  -I_1(\kappaa - g_\sigma)}{I_2} z^2e^{-\kappac z}
\end{multline}
\begin{equation}
  I_0 = \frac{4\pi h_\sigma (2 + \sigma \kappaa (2 + \sigma \kappaa))}{\kappaa^3}
  - \frac{4\pi}{3} \sigma^3
\end{equation}
\begin{equation}
  I_1 = \frac{4\pi h_\sigma (6 + \sigma \kappab (4 + \sigma \kappab))}{\kappab^4}
\end{equation}
\begin{equation}
  I_2 = \frac{8\pi h_\sigma (12 + \sigma \kappac (6 + \sigma \kappac))}{\kappac^5}
\end{equation}
\begin{equation}
  z = r - \sigma
\end{equation}
\davidsays{Can someone check the above equations for correctness
  against the python code, in plot-ghs2.py.  We'll want to
  ensure the C++ code agrees with this as well.}
This function includes three fitted parameters, $\kappaa$, $\kappab$,
and $\kappac$, which are given in Table~\ref{tab:fit-parameters}.
The fit is displayed in Fig.~\ref{fig:radial-distribution}.  In that
figure, it is apparent how the on the integral of $h(r)$ from
Eq.~\ref{eq:total-constraint} results in cancelling errors at large
distances.  The maximum error in $g(r)$ at packing fractions of less
than 0.4 is \fixme{get a number}.

\begin{table}
  \begin{tabular}{cc}
    $\kappaa$ & 1 \\
    $\kappab$ & 1 \\
    $\kappac$ & 1 \\
  \end{tabular}
  \caption{Fit parameters \fixme{... put them here and format this nicely}}
  \label{tab:fit-parameters}
\end{table}

%% In addition, we know that
%% \begin{align}
%%   1 + n\int\left[g(r)-1\right]d\rr &= nkT\chi_T
%% \end{align}
%% \begin{equation}
%%   \begin{split}
%%     \chi_T  &= -\frac{1}{V}\left(\frac{\partial V}{\partial
%%       p}\right)_T \\
%%     &= - \left(V\left(\frac{\partial p}{\partial
%%       V}\right)_T\right)^{-1}
%%   \end{split}
%% \end{equation}
%% \begin{equation}
%%     p = nkT\frac{1+\eta+\eta^2-\eta^3}{(1-\eta)^3}
%% \end{equation}
%% \begin{equation}
%%     \left(\frac{\partial p}{\partial V}\right)_T = \frac{\partial p}{\partial\eta}\frac{\partial\eta}{\partial V} + \frac{\partial p}{\partial n}\frac{\partial n}{\partial V}
%% \end{equation}
%% \begin{equation}
%%     n = \frac{N}{V}\Longrightarrow \frac{\partial n}{\partial V} = -\frac{N}{V^2} = -\frac{n}{V}
%% \end{equation}
%% \begin{equation}
%%     \eta = \frac{4\pi R^3N}{3V}\Longrightarrow \frac{d\eta}{dV} = -\frac{\eta}{V}
%% \end{equation}
%% \begin{equation}
%%   \begin{split}
%%     \left(\frac{\partial p}{\partial V}\right)_T &= -nkT\frac{\eta}{V}\left[\frac{1 + 2\eta - 3\eta^2}{(1-\eta)^3} + 3\frac{1 + \eta + \eta^2 - \eta^3}{(1-\eta)^4}\right]\\
%%     &\quad  - \frac{nkT}{V}\left[\frac{1+\eta+\eta^2-\eta^3}{(1-\eta)^3}\right]\\
%%     &= -nkT\frac{\eta}{V}\left[\frac{(1+2\eta-3\eta^2)(1-\eta) + 3 + 3\eta + 3\eta^2 - 3\eta^3}{(1-\eta)^4}\right]\\
%%     &\quad  - \frac{nkT}{V}\left[\frac{(1+\eta+\eta^2-\eta^3)(1-\eta)}{(1-\eta)^4}\right]\\
%%     &= -\frac{nkT}{V}\left[\frac{4\eta + 4\eta^2 - 2\eta^3}{(1-\eta)^4} + \frac{1 - 2\eta^3 + \eta^4}{(1-\eta)^4}\right]\\
%%   \end{split}
%% \end{equation}
%% \begin{equation}
%%     nkT\chi_T = \frac{(1-\eta)^4}{1 + 4\eta + 4\eta^2 - 4\eta^3 + \eta^4}
%%   %\int (g(r)-1)d\rr &= \frac{\chi_T}{kT} \\
%%   %&= \eta \eta \cdots \fixme{fill in, double-check}
%% \end{equation}
%% where $\chi_T$ is the isothermal compressibility, for which we use the
%% Carnahan-Starling equation of state.

%% An additional constraint that we use sets the slope of $g(r,
%% g_\sigma)$ at $r=\sigma$.  \fixme{explain this}

%% Given these constraints, we use a functional form of
%% \begin{equation}
%%   g(r,g_\sigma) = 1 + A e^{-a(r-\sigma)}
%%   + B (r-\sigma) e^{-b(r-\sigma)} + C (r-\sigma)^2 e^{-c(r-\sigma)}
%% \end{equation}
%% where $A$, $B$, and $C$ are amplitudes determined by the above
%% constraints, and $a$, $b$, and $c$ are parameters fit to Monte Carlo
%% simulations of the hard-sphere radial distribution function.
